---
title: "Exam 3 Skills Test"
author: "Joshua Nilsson"
date: "2026-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(janitor)
library(easystats)
library(broom)
```

## Question 1: Clean the table, and reconstruct the example graph.

In this section, I loaded in our data and converted the Salary columns into rows, the names of each column went to Rank and the values to Salary. I updated the professor titles so they were more clear. Then I dropped the unneeded information. 
```{r load-salaries}
salaries <- read.csv("FacultySalaries_1995.csv")
# Cleaning the csv to add the necessary rows for our graph.
clean_salaries <- salaries %>% 
  pivot_longer(cols = matches("Salary"),
               names_to = "Rank",
               values_to = "Salary") %>% 
  mutate(
    Rank = case_when(
      Rank == "AvgFullProfSalary"   ~ "Full",
      Rank == "AvgAssocProfSalary"  ~ "Associate",
      Rank == "AvgAssistProfSalary" ~ "Assistant",
      Rank == "AvgProfSalaryAll"    ~ "All",
      TRUE ~ Rank
    )
  ) %>% 
  filter(Rank != "All", Tier != "VIIB")  #Excluding the "All" group and "VIIB" tier
```
I used a facet_wrap() to separate them by Tier, and theme() to alter the x-axis titles.
```{r plot-salaries, message=FALSE, warning=FALSE}
# Plotting our graph.
plot <- clean_salaries %>% 
  ggplot(aes(x = Rank, y = Salary, fill = Rank)) +
  geom_boxplot() +
  facet_wrap(~ Tier) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotating the x-axis text.

plot
```

## Question 2: Build an ANOVA model that takes into consideration "State”, “Tier”, “Rank” on “Salary” 

I used the aov() function to create the ANOVA model, then I summarized the model using summary(). 
```{r anova-model}
aovmod <- clean_salaries %>% 
  aov(formula = Salary ~ State + Tier + Rank)
summary(aovmod) #Prints the summary output
```

## Question 3: Tidy the “Juniper_Oils.csv” data

To clean this, I took the various chemical columns and turned them into rows. The names became the ChemicalID, and the values became their concentration.  
```{r load-oils}
oils <- read.csv("Juniper_Oils.csv", check.names = FALSE)
clean_oils <- oils %>% 
  pivot_longer(cols = matches(c("alpha-pinene","para-cymene","alpha-terpineol","cedr-9-ene","alpha-cedrene","beta-cedrene","cis-thujopsene","alpha-himachalene","beta-chamigrene","cuparene","compound 1","alpha-chamigrene","widdrol","cedrol","beta-acorenol","alpha-acorenol","gamma-eudesmol","beta-eudesmol","alpha-eudesmol","cedr-8-en-13-ol","cedr-8-en-15-ol","compound 2","thujopsenal")),
               names_to = "ChemicalID",
               values_to = "Concentration")
```

## Question 4: Make a graph of Chemical Concentrations through time

For graphing, I used the ggplot function, and I made sure the proper inputs were in facet_wrap() function to match the example graph provided. 
```{r plot-oils, message=FALSE, warning=FALSE}
plot2 <- clean_oils %>% 
  ggplot(aes(x = YearsSinceBurn, y = Concentration)) +
  facet_wrap(~ChemicalID, axes = "all_x", scales = "free_y") +
  geom_smooth()

plot2
```

## Question 5: Create a GLM to find which chemicals are significantly affected by "Years Since Burn"

I used the glm() function to make the model, and I made sure the interactions between the ChemicalID and YearsSinceBurn were included to find the significant chemicals within the model. From there it was just a matter of cleaning the output term names and making sure it only displayed parameters that were below the 0.05 p-value threshold.
```{r glm-model, message=FALSE, warning=FALSE}
glmmod <- glm(data = clean_oils, formula = Concentration ~ ChemicalID + ChemicalID:YearsSinceBurn)

glmmod %>% 
  tidy() %>% 
  filter(p.value < 0.05) %>% 
  mutate(term = gsub("ChemicalID", "", term))
```

